---
title: "COVID-19 Application"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    source_code: 
    theme: simplex

---

```{r setup, include=FALSE}
library(shiny)
library(leaflet) 
library(tidyverse)
library(plotly) 
library(ggplot2) 
library(sp)
library(dplyr)
library(flexdashboard) 
library(maptools)
library(readr)
library(viridis)
library(mapview)
library(leafsync)
library(geojsonio)
options(scipen = 999)

# Define the list of available metros
lookupcountry <- structure(c("Sweden", "Afghanistan","Germany", "France", "United Kingdom"))
lookupcountry2 <- structure(c("Sweden", "Afghanistan", "Germany", "France", "United Kingdom"))
lookupinterest <- structure(c("Confirmed cases", "Deaths", "Recovered"))
lookupscale <- structure(c("Natural log", "Raw scale"))
lookupsmoother <- structure(c("No", "Yes"))
lookupstart <- structure(c("Calendar date", "Time since first case"))

# Read in data, and subset for the selected metro
confirmed <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
deaths <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv") 
recovered <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv") 

#Worldcountry
WorldCountry <-geojsonio::geojson_read("https://raw.githubusercontent.com/johan/world.geo.json/master/countries.geo.json", what = "sp")

WorldCountry$id[which(WorldCountry$id == "SWE")] <- "Sweden"
WorldCountry$id[which(WorldCountry$id == "AFG")] <- "Afghanistan"
WorldCountry$id[which(WorldCountry$id == "DEU")] <- "Germany"
WorldCountry$id[which(WorldCountry$id == "FRA")] <- "France"
WorldCountry$id[which(WorldCountry$id == "GBR")] <- "United Kingdom"
WorldCountry$id[which(WorldCountry$id == "BRA")] <- "Brazil"
WorldCountry$id[which(WorldCountry$id == "USA")] <- "US"

#Write a function for taking the dateframe
#Deaths
deaths <- deaths %>% 
  gather(Date, Outcome, 5:length(deaths)) %>%
  rename(Country = `Country/Region`) %>%
  rename(Province = `Province/State`)

deaths <- deaths %>%
  mutate(Date = lubridate::mdy(Date))

#Confirmed
confirmed <- confirmed %>% 
  gather(Date, Outcome, 5:length(confirmed)) %>%
  rename(Country = `Country/Region`) %>%
  rename(Province = `Province/State`)

confirmed <- confirmed %>%
  mutate(Date = lubridate::mdy(Date))

#Recovered
recovered <- recovered %>% 
  gather(Date, Outcome, 5:length(recovered)) %>%
  rename(Country = `Country/Region`) %>%
  rename(Province = `Province/State`)

recovered <- recovered %>%
  mutate(Date = lubridate::mdy(Date))

x <- list(confirmed, deaths, recovered)

m <- reactive({
  if(input$outcome == "Confirmed cases"){
    m <- as.data.frame(x[1])
    m <- m %>% filter((m$Country %in% input$country3 & is.na(m$Province)))
    if(input$start == "Time since first case"){
    m <- m %>% filter(m$Outcome > 0)  
    }
    return(m)
  }
  else if(input$outcome == "Deaths"){
    m <- as.data.frame(x[2])
    m <- m %>% filter((m$Country %in% input$country3 & is.na(m$Province))) 
    if(input$start == "Time since first case"){
    m <- m %>% filter(m$Outcome > 0)  
    }
    return(m)
  }
  else {
    m <- as.data.frame(x[3])
    m <- m %>% filter((m$Country %in% input$country3 & is.na(m$Province)))
    if(input$start == "Time since first case"){
    m <- m %>% filter(m$Outcome > 0)  
    }
    return(m)
  }
  
})
```

Sidebar {.sidebar}
======================================================================

```{r}
checkboxGroupInput("country3", "Countries to show:",
                           c("Sweden" = "Sweden",
                             "Afghanistan" = "Afghanistan",
                             "Germany" = "Germany",
                             "France" = "France",
                             "United Kingdom" = "United Kingdom",
                             "Brazil" = "Brazil",
                             "USA" = "US"))

output$txt <- renderText({
    country3 <- paste(input$country3, collapse = ", ")
    paste("You chose", country3)
  })

#Define country
#selectInput('country', label = 'Select a Country', choices = lookupcountry, selected = "Sweden")

#Define country
#selectInput('country2', label = 'Select a Country for comparison', choices = lookupcountry, selected = "Afghanistan")

#Define natural log or raw scale 
selectInput('scale', label = 'Select a scale', choices = lookupscale, selected = "Raw scale")
#Define Smoothed
#selectInput('smoother', label = 'Smoother', choices = lookupsmoother, selected = "No")
checkboxInput("smoother", "Add smoother", FALSE)
#Define outcome
selectInput('outcome', label = 'Select your outcomes of interest', choices = lookupinterest, selected = "Confirmed cases")
#Define start
selectInput('start', label = 'Start by calendar date or time since first case', choices = lookupstart, selected = "Calendar date")

```

Use the __Explore country trends__ tab to explore time series trends for your chosen country. 

Click the __Compare countries__ tab to perform a comparative evaluation between countries for the selected variables of interest. To learn more about the project, click the __About__ tab.  

Application author: [Tommy Ivanics](https://github.com/tivanics), [Johns Hopkins Bloomberg School of Public Health](https://www.jhsph.edu)

Explore country trends
======================================================================

Row
-----------------------------------------------------------------------

### Time series trends in selected country

```{r}
output$scatter <- renderPlotly({
  if(input$scale == "Raw scale"){
  p1a <- ggplot(m()) + 
    geom_point(alpha = 0.4, aes(Date, Outcome, fill=Country)) + 
    theme_minimal(base_size = 14) + 
    labs(x="Date", y=input$outcome)
  if(input$smoother == 1){
    for(i in seq_along(input$country3)){
    p1a <- p1a + stat_smooth(data=subset(m(), Country %in% input$country3[i]), 
                aes(Date, Outcome), 
                color = '#FFD700', 
                method = 'loess',
                se = FALSE,
                alpha = 1)
    i+1
    }
    p1a
  }
  else{
    p1a
  }
  }
  else if(input$scale == "Natural log"){
  p1a <- ggplot(m()) + 
    geom_point(alpha = 0.4, aes(Date, log(Outcome), fill=Country)) + 
    theme_minimal(base_size = 14) + 
    labs(x="Date", y=paste("Natural log of", input$outcome))
  if(input$smoother == 1){
    for(i in seq_along(input$country3)){
        p1a <- p1a + stat_smooth(data=subset(m(), Country %in% input$country3[i]), 
                aes(Date, log(Outcome)), 
                color = '#FFD700', 
                method = 'loess',
                se = FALSE,
                alpha = 1)
        i+1
    }
    p1a
  }
  else{
    p1a
  }
  }
})  

plotlyOutput('scatter', width = "80%")
```

Row 
-----------------------------------------------------------------------
### Map of selected country

```{r}
output$maps <- renderLeaflet({
    data_Map <- WorldCountry[WorldCountry$id %in% m()$Country, ]
map <- leaflet(data_Map) %>% addTiles() %>% addPolygons()
    
})  
leafletOutput('maps', width = "80%")
```

About
============================================================================

This application is for the course **Data Science for Public Health**, ["Data Science in Biomedical Engineering and Data Science for Public Health."](https://github.com/bcaffo/ds4ph-bme).

The COVID-19 application analyzes international temporal trends in case number, deaths, and recovered cases.

There are three key metric in this article: **Case numbers** **Deaths** **Recoveries** 

This application allows users to explore these outcomes interactively.  The user can select a scale **Natural log or raw scale**, choose to add a **loess smoother**, and select a start time by **calendar date** or **date of first case**. 

Temporal (time-series) data for the application come from [COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University](https://github.com/CSSEGISandData/COVID-19), and individual data for outcomes can be found here: [Cases](https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv) [Deaths](https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv) and [Recoveries](https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv).

The application is built with the [Shiny](http://shiny.rstudio.com) framework for the [R programming language](https://www.r-project.org/). The application layout is produced with the [flexdashboard](http://rstudio.github.io/flexdashboard/index.html) package, and the charts and maps use [Plotly](http://plot.ly), [Leaflet.js](http://leafletjs.com/), and [ggplot2](http://ggplot2.org/), all accessed through their corresponding R packages.  

I welcome feedback and suggestions!  [Please visit my GitHub](https://github.com/tivanics/) for contact information or [connect with me on Twitter](https://twitter.com/ivanics_t).  


<style>

.section.sidebar {

  background-color: white; 
  font-family: "Open-Sans", "Helvetica Neue", Helvetica, Arial, sans-serif;

}

.js-irs-0 .irs-bar {
border-top-color: #443A83;
border-bottom-color: #443A83;
} 

.js-irs-0 .irs-bar-edge {
border-color: #443A83;
}

.js-irs-0 .irs-single, .js-irs-0 .irs-bar-edge, .js-irs-0 .irs-bar {
background: #443A83;
}

.navbar-inverse {
background-color: #443A83;
border-color: #440154;
}

.navbar-inverse .navbar-brand {
color: #a3a9ac;
}

a:hover, a:focus {
color: #440154;
text-decoration: underline;
}

a {
color: #443A83;
text-decoration: none;
}

.navbar-inverse .navbar-nav>li>a {
color: #a3a9ac;
}

</style>




